Analysing Constraint Grammar with a SAT solver

### 2015-10-13 paper/lrec/problem_desc.tex
add quote by atro
move text around

As CGs grow larger, it gets harder for the grammar writers and users to keep track of all the rules and their interaction.
\cite{bick2013tuning} presents efforts to optimise hand-written CGs using machine learning techniques.
While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
Our toolset is designed to complement the existing approaches in CG analysis and optimisation.
The ideal use case would be a grammar writer, who wants to know e.g.

@@

Our work draws inspiration from various sources. This section explains what our approach is and is not, and relates it to the previous research. We combine elements from the following aspects of CG research:

\begin{itemize}
\item Methods to aid in hand-crafting rules \cite{voutilainen2004}, \cite{tyers_reynolds2015}
\item Optimising a hand-written CG \cite{bick2013tuning}
\item Encoding CG in logic \cite{lager98}, \cite{lager_nivre01}, \cite{lager01transformation}, \cite{listenmaa_claessen2015}
\end{itemize}

There is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996}, \cite{lindberg_eineborg98ilp}%, \cite{asfrent14}
. Our work is aimed to aid the process of hand-crafting rules.

@@

Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin \cite{voutilainen2004}. For a language with no free hand-tagged corpus available, \cite{tyers_reynolds2015} describe a method where they apply their rules to Wikipedia texts and pick 100 examples at random for manual check.
\begin{itemize}
\item \# times the rule was applied
\item \% times it was correct
\end{itemize}
Strengths: gives concrete feedback about the effect of rules. 
Good especially at the beginning stages of grammar writing. 
Sometimes you just want to know that a rule fired 500 times and was correct 490 times, rather than ``it is possible to construct a word sequence that will trigger this rule''.
The connection to the reality has also weak sides. This method will not notice a missed opportunity. What if instead of firing 500 times, it could have fired 1000 times if it had been placed earlier in the rule sequence.
Or the rule that was wrong in 60 \% of the time,

% \cite{voutilainen2004} state that the first 200 rules are probably enough to resolve 50--75 \% of ambiguities in the corpus used in the development. 
Authors writing grammars for well-resourced languages can also benefit from the technique, as an additional tool along with a gold standard corpus.

\paragraph{ML optimisation of hand-written grammars}
\begin{itemize}
\item Variations in the place of the rule
\item Variations in the conditions (careful, ...) of the rule
\end{itemize}
Good side: try all combinations to find the best result. 
At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions. \cite{bick2013tuning} tries out combinations of moving rules in different sections or removing them in total, and in parallel, making their contexts stricter or less strict. 
This is a valuable tool, especially for grammars that are so big that it's hard to keep track of. Program can try all combinations whereas trying to make sense out of a huge set of rules would be hard for humans.
As a downside, the grammar writer will likely not know why exactly is the tuned grammar performing better.

>>>>

As CGs grow larger, it gets harder for the grammar writers and users to keep track of all the rules and their interaction.
\cite{voutilainen2004} gives a detailed account about best practices of grammar writing, and efficient use of hand-annotated corpora to aid the grammar development.
\cite{bick2013tuning} presents a method to optimise hand-written CGs using machine learning techniques.
Our toolset is designed to complement the existing approaches in CG analysis and optimisation.
While Bick's experiments show improvements in results, the grammar writer is none the wiser why is the grammar better.
Authors who have a large hand-annotated corpus can also benefit from the technique.
With our method, it is possible to ask, for example, the following questions:

@@

\begin{quote}Another desirable facility in the grammar development environment would
be a mechanism for identifying pairs of constraints that contradict each
other.
--Atro Voutilainen, 2004
\end{quote}

This section describes prior research and relates our contribution to the existing approaches.
We combine elements from the following aspects of CG research:

\begin{itemize}
\item Corpus-based methods in manual grammar development \cite{voutilainen2004}, \cite{tyers_reynolds2015}
\item Optimising hand-written grammars \cite{bick2013tuning}
\item Encoding CG in logic \cite{lager98}, \cite{lager_nivre01}, \cite{listenmaa_claessen2015}
\end{itemize}

In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996}, \cite{lindberg_eineborg98ilp}%, \cite{lager01transformation}, \cite{asfrent14} 
. However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.

@@

Hand-annotated corpora are commonly used in the development of CGs, because they give immediate feedback whether a new rule increases or decreases accuracy, and by how large margin \cite{voutilainen2004}. For a language with no free or tagset-compatible corpus available, \cite{tyers_reynolds2015} describe a method where they apply their rules to unannotated Wikipedia texts and pick 100 examples at random for manual check.

% \begin{itemize}
% \item \# times the rule was applied
% \item \% times it was correct
% \end{itemize}
CG rules are usually arranged in sections, from safest to least safe. Given a grammar that has three sections, the rules are applied to the text in the following manner. First apply rules from section 1, and repeat until nothing changes in the text. Then apply rules from sections 1--2, and finally all rules, from sections 1--3.
The best strategy is to start with the most effective rules with low error rate, so that they make way for the following, more heuristic and less safe rules to act on.

A representative corpus is arguably the best way to get concrete numbers: how many times a rule applied and how often it was correct.
However, this method will not notice a missed opportunity, nor suggest a way to improve, for instance changing the place of a rule.

% \cite{voutilainen2004} state that the around 200 rules are probably enough to resolve 50--75 \% of ambiguities in the corpus used in the development. 
% This is invaluable, especially at the beginning stages of grammar writing. 
% Sometimes you just want to know that a rule fired 500 times and was correct 490 times, rather than ``it is possible to construct a word sequence that will trigger this rule''.

The connection to the reality has also its limitations. 
% What if instead of firing 500 times, it could have fired 1000 times if it had been placed earlier in the rule sequence.
Or the rule that was wrong in 60 \% of the time,

\paragraph{Optimising hand-written grammars}
\begin{itemize}
\item Variations in the place of the rule
\item Variations in the conditions (careful, ...) of the rule
\end{itemize}
Good side: try all combinations to find the best result. 
At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions. \cite{bick2013tuning} tries out combinations of moving rules in different sections or removing them in total, and in parallel, making their contexts stricter or less strict. 
This is a valuable tool, especially for grammars that are so big that it's hard to keep track of. Program can try all combinations whereas trying to make sense out of a huge set of rules would be hard for humans.
As a downside, the grammar writer will likely not know why exactly is the tuned grammar performing better.
